Linked Open Social Data for Scientific Benchmarking

### 2016-07-19 paper.tex
advances for first preliminary version

%% `Elsevier LaTeX' style
\bibliographystyle{elsarticle-num}
%
\makeatletter
\def\ps@pprintTitle{%
    \let\@oddhead\@empty
    \let\@evenhead\@empty
    \def\@oddfoot{}%
\let\@evenfoot\@oddfoot}
\makeatother
\begin{document}
%
\begin{frontmatter}
%
\title{Linked Open Social Data for Scientific Benchmarking}
%
\author[pwr]{Renato Fabbri\corref{corresponding}\fnref{kio-url}}
\ead{fabbri@usp.br}
%
\author[pwr]{Osvaldo Novais de Oliveira Junior\fnref{kio-url}}
\ead{chu@ifsc.usp.br}
%
\cortext[corresponding]{Corresponding author}
\address[pwr]{S\~ao Carlos Institute of Physics, S\~ao Paulo
University, Brazil}
%
\fntext[kio-url]{\textit{URL:} \url{http://www.ifsc.usp.br/}}
%
\begin{abstract}
The field of social network analysis and the topic of complex networks
are widely researched.
Recently, a myriad of results have been reported which are based in
diverse datasets most often not accessible to other researchers.
This work exposes an open dataset with diverse provenance and oriented
to provide the scientific community a friendly and common repertoire.
Current data was obtained from Facebook, Twitter, IRC, Email and the
detached instances of ParticipaBR, AA and Cidade Democr\'atica.
These were represented as linked data to homonenize access,
conform to current best practices and ease analyzes which integrate third
party and provided instances.
This document presents an outline and overall statistics of the given
dataset which should favor subsequent work.
\end{abstract}
%
\begin{keyword}
Big Data, Data Mining, Benchmark Data, Facebook, Twitter, IRC, Email, Complex Networks
%Hierarchy of Clusters \sep HoC \sep Benchmark Dataset \sep Benchmark Data Generator \sep Artificial Data \sep Cluster Analysis \sep Tree Structured Stick Breaking Process \sep TSSB \sep ...
\end{keyword}

@@

\section{Introduction}
In recent years, the web of linked data~\cite{lee1} has attracted wide attention in
both research and application realms.
However, there is is a lack of datasets for research benchmarking,
specially in the complex networks field, yielding diverse results from 
poorly related data.

The enormity of the digital data propels a rapid development of analysis methods
from different perspectives.
The used datasets differ within the scope of each research with
scarce and historical exception such as the
karate club dataset~\cite{newmanBook}.
On the other hand, the available linked data is not 
stable or rigorous enough to be
a public reference on statistical physics and social networks research.

This work presents a linked open social data (LOSD) dataset with data from diverse
provenance, including Facebook, Twitter, IRC, Email and detached
instances.
Such data is proposed as a common repertoire for scientific
reseach involving networks and textual content.

@@

\section{Materials}\label{materials}
Data was gathered either from:
\begin{itemize}
    \item public APIs (Twitter, Email); or
    \item public logs (IRC and AA); or
    \item Netvizz software~\cite{netvizz} (Facebook); or
    \item donated data from users (Facebook); or
    \item donated data from system administrators (AA, ParticipaBR,
        Cidade Democr\'atica).
\end{itemize}

@@

\section{Results}
\label{outline}
Knowledge Discovery in Databases finished untill data preparation phase.
\subsection{Standardization}
Standards: snapshot ids: <platform>-legacy-<further\_identifier>;
ex: irc-legacy-labmacambira; email-legacy-c++;
facebook-legacy-RenatoFabbri, etc., participants and transactions linkage to
snapshots.

Participation ontology URIs. Concept URIs are the same for different
snapshot provenance, being differetiable by the snapshot metadata.

Big data.
\subsection{Data outline}
Relations only (Facebook), text only (AA, Cidade Democratica).
\input{misc/overallText}
\input{tables/nsnapshots}
\input{tables/basicOverall}
% stats:
%% number of triples
%% number of edges
%% number of chars
%% number of users
% diagrams in the supporting information file
\subsection{Software tools}
The LOSD is released with a software for rendering itself, analyses and
multimedia artifacts.
\subsubsection{Triplification routines}
For each social platform there is a \emph{triplification} routine,
i.e. a script for translating data to RDF.
Original formats and further observations are presented in
Table~\ref{tab:provenance}.
\begin{table*}[h!]\scriptsize
\begin{center}
\caption{Social platforms, original formats and further observations for
the LOSD dataset.}\label{tab:provenance}
\begin{tabular}{| l || p{3cm} | p{3cm} | c |}\hline
    \textbf{social platform} & \textbf{original format} & \textbf{further observations} & \textbf{toolbox} \\\hline\hline
    AA & MySQL and MongoDB databases; IRC text logs & donated by AA users & Participation \\\hline
    Cidade Democrática & MySQL database & donated by admins & Participation \\\hline
    Email & mbox & obtained through Gmane public database & Gmane \\\hline
    Facebook & GDF, GML and TAB & obtained through Netvizz~\cite{netvizz} & Social \\\hline
    IRC & plain text log & obtained through Supybot logging & Social \\\hline
    ParticipaBR & PostgreSQL database & donated by admins & Participation \\\hline
    Twitter & JSON & obtained through Twitter streaming API & Social \\\hline
\end{tabular}\end{center}
\end{table*}                    
\subsubsection{Topological and textual Analysis}\label{ana}
Routines are available for taking topological and textual measures from
the dataset.
Principal Component Analysis (PCA) is performed with available measures
to ease pattern recognition both in a timeline and in a multiscale
fashion.

@@

\subsection{SPARQL queries}\label{queries}
% all text from user
% users with the most friendships
% users with most interactions
% search string in messages

>>>>

%% `Elsevier LaTeX' style
\bibliographystyle{elsarticle-num}
%
\makeatletter
\def\ps@pprintTitle{%
    \let\@oddhead\@empty
    \let\@evenhead\@empty
    \def\@oddfoot{}%
\let\@evenfoot\@oddfoot}
\makeatother
\begin{document}
%
\begin{frontmatter}
%
\title{Linked Open Social Data for Scientific Benchmarking}
%
\author[pwr]{Renato Fabbri\corref{corresponding}\fnref{kio-url}}
\ead{fabbri@usp.br}
%
\author[pwr]{Osvaldo Novais de Oliveira Junior\fnref{kio-url}}
\ead{chu@ifsc.usp.br}
%
\cortext[corresponding]{Corresponding author}
\address[pwr]{S\~ao Carlos Institute of Physics, S\~ao Paulo
University, Brazil}
%
\fntext[kio-url]{\textit{URL:} \url{http://www.ifsc.usp.br/}}
%
\begin{abstract}
The field of social network analysis and the topic of complex networks
are widely researched.
Recently, a myriad of results have been reported which are based in
diverse datasets most often not accessible to other researchers.
This work exposes an open dataset with diverse provenance and oriented
to furnish the scientific community with a friendly and common repertoire.
Current data was obtained from Facebook, Twitter, IRC, Email and the
detached instances of ParticipaBR, AA and Cidade Democr\'atica.
These were represented as linked data to homonenize access,
conform to current best practices and ease analyzes which integrate third
party and provided instances.
This document presents an outline and overall statistics of the given
dataset which should favor subsequent work.
\end{abstract}
%
\begin{keyword}
Big Data, Data Mining, Benchmark Data, Facebook, Twitter, IRC, Email, Complex Networks
%Hierarchy of Clusters \sep HoC \sep Benchmark Dataset \sep Benchmark Data Generator \sep Artificial Data \sep Cluster Analysis \sep Tree Structured Stick Breaking Process \sep TSSB \sep ...
\end{keyword}

@@

\section{Introduction}
In recent years, the web of linked data~\cite{lee1} has attracted wide attention in
both research and application realms.
However, there is a lack of datasets for research benchmarking,
specially in the complex networks field, yielding diverse results from 
poorly related data.

The enormity of the digital data propels a rapid development of analysis methods
from different perspectives.
However, the used datasets in the literature differ within the scope of each research with
scarce and historical exceptions such as the
karate club dataset~\cite{newmanBook}.
On the other hand, the available linked data is not 
stable or rigorous enough to be
a public reference on statistical physics and social networks research.

This work presents a linked open social data (LOSD) dataset with diverse
provenance, including Facebook, Twitter, IRC, Email and detached
instances.
Such data is proposed as a common repertoire for scientific
reseach involving networks and textual content.

@@

\section{Materials}\label{materials}
Data was gathered either from:
\begin{itemize}
    \item public APIs (Twitter, Email); or
    \item public logs (IRC and AA); or
    \item Netvizz software~\cite{netvizz} (Facebook); or
    \item donated data by users (Facebook); or
    \item donated data by system administrators (AA, ParticipaBR,
        Cidade Democr\'atica).
\end{itemize}

@@

\section{Results}
\label{outline}
Current overall results concern data selection and preparation for knowledge discovery.
The main result is the data made available, which enables benchmarking of scientific results
and easy experimentations.
Secondary results include software support and SparQL queries made beforehand.

\subsection{Standardization}
The LOSD is a collection of data translated into RDF and embedded into standard URIs and triples.
Such standards constitute the LOSD itself, but some of the conventions are outlined in this section.

URIs have the root \url{http://purl.org/socialparticipation/participationontology/} which are identified with
the radical \textttt{po:}.
Classes and properties are build by adding a suffix to the root, as in \textttt{po:Participant} or \textttt{po:text}.
Classes have ``UpperCamelCase'' suffixes while properties have ``lowerCamelCase'' suffixes.
All participants and transactions (messages, friendships, interactions) are linked to
snapshots through the triple \textttt{<resource> po:snapshot <snapshot\_uri>}.
Message texts, including comments, are objectives in the triple: \textttt{<message\_id> po:text <message\_text>}.
Individuals (also called instances) are built on top of the class they derive from plus a hashtag character,
the snapshot id of the snapshot they refer to, and an identifier;
i.e. \textttt{po:Participant\#<snapshot\_id>-<local\_id>}.
All snapshot URIs follow the formation rule: \textttt{po:<SnapshotProvenance>\#<snapshot\_id>}.
All snapshot ids follow the formation rule: \textttt{<platform>-legacy-<further\_identifier>}; e.g.
\textttt{irc-legacy-labmacambira} or \textttt{email-legacy-c++...}.

\subsection{Data outline}
\input{misc/overallText}

\input{tables/basicOverall}

\input{tables/nsnapshots}
% stats:
%% number of triples
%% number of edges
%% number of chars
%% number of users
% diagrams in the supporting information file
\subsection{Software tools}
The LOSD is released with a software for rendering itself, analyses and
multimedia artifacts.
\subsubsection{Triplification routines}
For each social platform there is a \emph{triplification} routine,
i.e. a script for translating data to RDF.
Original formats and further observations are presented in
Table~\ref{tab:provenance}.
\begin{table*}[h!]\scriptsize
\begin{center}
\caption{Social platforms, original formats and further observations for
the LOSD dataset.}\label{tab:provenance}
\begin{tabular}{| l || p{3cm} | p{3cm} | c |}\hline
    \textbf{social platform} & \textbf{original format} & \textbf{further observations} & \textbf{toolbox} \\\hline\hline
    AA & MySQL and MongoDB databases; IRC text logs & donated by AA users & Participation \\\hline
    Cidade Democrática & MySQL database & donated by admins & Participation \\\hline
    Email & mbox & obtained through Gmane public database & Gmane \\\hline
    Facebook & GDF, GML and TAB & obtained through Netvizz~\cite{netvizz} & Social \\\hline
    IRC & plain text log & obtained through Supybot logging & Social \\\hline
    ParticipaBR & PostgreSQL database & donated by admins & Participation \\\hline
    Twitter & JSON & obtained through Twitter streaming API & Social \\\hline
\end{tabular}\end{center}
\end{table*}                    
\subsubsection{Topological and textual Analysis}\label{ana}
Routines are available for taking topological and textual measures from
the dataset.
Principal Component Analysis (PCA) is performed with available measures
to ease pattern recognition both in a timeline and in a multiscale
fashion.

@@

\subsection{SPARQL queries}\label{queries}
There are numerous useful and general purpose SPARQL queries to be performed agaist the LOSD database.
Here we write some of the most basic of such queries selected by their potential to be varied.
All queries assume the use the preable \textttt{PREFIX po: <http://purl.org/socialparticipation/po/>}.
\begin{enumerate}[leftmargin=0cm]
	\item Retrieve the number of participants
	\item Retrieve the number of relations, be them interactions of friendships
	\item Retrieve the number of relations, be them interactions of friendships
	\item Retrieve all text produced by an specific user
	\item List 100 users with the most friendships
	\item List 100 users with most interactions
	\item Search string in messages
\end{enumerate}
